---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# CTRE

This R package provides tools to analyse extremes of 'bursty' time series. 
[Burstiness](https://en.wikipedia.org/wiki/Burstiness) is characterized by 
heavy-tailed inter-arrival times and scale-free event dynamics. 
The CTRE model captures burstiness by generalizing the Poisson process 
to a 
[fractional Poisson process](https://en.wikipedia.org/wiki/Fractional_Poisson_process), 
with [Mittag-Leffler](https://strakaps.github.io/MittagLeffleR/) 
inter-arrival times. 
Parameter estimates are read off from stability plots, and 
goodness of fit is assessed via diagnostic plots; see the 
Shiny app below. 

## Shiny App

The package comes with two examples of bursty time series: solar flare 
magnitudes and bitcoin trading volumes. 
For parameter estimates of the Mittag-Leffler distribution, see the tab 
"Exceedance Times". 
CTRE model assumptions are checked via a QQ plot of the Mittag-Leffler
distribution; an empirical copula plot checking for dependence between 
inter-arrival times and magnitudes; and a plot of the autocorrelation 
function for the two series (interarrival times and magnitudes). 
For the standard POT model plots, see the "Exceedances" tab. 


<div>
<iframe src="https://strakaps.shinyapps.io/ctre-app/" style="border: none; width: 800px; height: 900px"></iframe>
</div>

## Install from GitHub

```{r, eval=FALSE}
library("devtools")
install_github("UNSW-MATH/CTRE")
library(CTRE)
```

## Run shiny app

You can run the above Shiny app from within RStudio: 

```{r, eval=FALSE}
runCTREshiny()
```


## Package usage

You can 

* Create a [`ctre`](/CTRE/reference/ctre.html) object from a time series, 
  a data frame, or two vectors.
* Plot it with [`plot`](/CTRE/reference/plot.ctre.html)
* Discard the data below a threshold with [`thin`](/CTRE/reference/thin.html)
* Extract data with [`interarrival`](/CTRE/reference/interarrival.html), 
  [`time`](/CTRE/reference/time.html) and 
  [`magnitudes`](/CTRE/reference/magnitudes.html)
* Create stability plots with [`MLestimates`](/CTRE/reference/MLestimates.html)
* Look at diagnostic plots 
  ([`mlqqplot`](/CTRE/reference/mlqqplot.html), [`acf`](/CTRE/reference/acf.html), 
  [`empcopula`](/CTRE/reference/empcopula.html))

