---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# CTRE

[![Travis-CI Build Status](https://api.travis-ci.org/UNSW-MATH/CTRE.svg?branch=master)](https://travis-ci.org/UNSW-MATH/CTRE)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/CTRE)](https://cran.r-project.org/package=CTRE)
[![Downloads](http://cranlogs.r-pkg.org/badges/CTRE)](https://cran.r-project.org/package=CTRE)
[![DL_Total](http://cranlogs.r-pkg.org/badges/grand-total/CTRE?color=blue)](https://cran.r-project.org/package=CTRE)

The CTRE package fits a CTRE model ("Continuous Time Random Exceedances") 
to extremes of 'bursty' time series. 

+ A [bursty](https://en.wikipedia.org/wiki/Burstiness) time series typically
  has heavy-tailed inter-arrival times.
+ Given a high threshold, the times between threshold crossings follow a 
  [Mittag-Leffler](https://strakaps.github.io/MittagLeffleR/) distribution, 
  whose scale parameter increases with the threshold.

The magnitudes of the extremes are modelled via the 
[POT](https://en.wikipedia.org/wiki/Extreme_value_theory#Data_analysis)
("Peaks-Over-Threshold") method, which is standard in Extreme Value Theory. 
It is the time between threshold crossings that is the focus of the CTRE model.

## Installation

### Stable release on CRAN

You can install `CTRE` from CRAN via 

```{r CRAN-installation, eval = FALSE}
install.packages("CTRE")
library(CTRE)
```

### Development version on Github

Install the `devtools` package first, then 

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("strakaps/CTRE")
library("CTRE")
```

## Usage

Below is the main workflow for fitting a CTRE model to a time series. 
See the [Reference page](/CTRE/reference/index.html) for more details on each 
R function. 

* Create a [`ctre`](/CTRE/reference/ctre.html) object from a time series, 
  a data frame, or two vectors.
* Plot it with [`plot`](/CTRE/reference/plot.ctre.html)
* Discard the data below a threshold with [`thin`](/CTRE/reference/thin.html)
* Extract data with [`interarrival`](/CTRE/reference/interarrival.html), 
  [`time`](/CTRE/reference/time.html) and 
  [`magnitudes`](/CTRE/reference/magnitudes.html)
* Create stability plots with [`MLestimates`](/CTRE/reference/MLestimates.html)
* Look at diagnostic plots 
  ([`mlqqplot`](/CTRE/reference/mlqqplot.html), [`acf`](/CTRE/reference/acf.html), 
  [`empcopula`](/CTRE/reference/empcopula.html))


## Shiny App

+ The package comes with two examples of bursty time series: solar flare 
magnitudes and bitcoin trading volumes. 
+ For parameter estimates of the Mittag-Leffler distribution, see the tab 
"Exceedance Times". 
+ For the POT model fit, see the tab "Exceedances"
+ CTRE model assumptions are checked via 
    - a QQ plot of the Mittag-Leffler distribution
    - an empirical copula plot checking for dependence between 
      inter-arrival times and magnitudes
    - a plot of the autocorrelation function for the two series (interarrival 
      times and magnitudes). 

To run the Shiny app from within RStudio:

```{r, eval=FALSE}
runCTREshiny()
```

<div>
<iframe src="https://strakaps.shinyapps.io/ctre-app/" style="border: none; width: 800px; height: 900px"></iframe>
</div>



## The CTRE paper

"Peaks Over Threshold for Bursty Time Series", 
Katharina Hees, Smarak Nayak, Peter Straka (2018). 
<https://arxiv.org/abs/1802.05218>
